---
title: "Heatmaps in Plotly"
output: html_notebook
---
Packages installed/opened:
  dplyr: I just opened this one in case I want to manipulate my table
  Tidyverse: it hosts the excel importer I used that doesn't require java
    readxl is that package
  plotly: the guest of honor
  


First, we'll want to open the package that will allow you to import Excel files
```{r}
library(dplyr)
library(tidyverse)
install.packages("readxl")
  #this is an alternative to the xlsx package that has no outside dependencies such as JAVA
library("readxl")
read_xlsx("TeamRed_Transposed_D10.xlsx")
  #this is the function with which you'll select the excel file. As long as it is in your working director, it will find it.
data = read_xlsx("TeamRed_Transposed_D10.xlsx")
  #that file name is just too long. We'll call it data. Very creative, I know.
```
 A heatmap will be perfect to give us visualization of the fold changes across the various conditions.

First, let's remind ourselves what our columns are.
```{r}
names(data) 
  #this will remind us of what the column names are

```
Now that I've been reminded, I want to double check that R has correctly determined their classes and rename anything that's not right.
```{r}
data=rename(data, D10="Day 10 Ratio to Day 0")
  #That column name was too much, so now it's called D10
class(data$Treatment)
  #should be "character"
class(data$Cytokine)
  #should be "character"
class(data$D10)
  #should be numeric


```

Now, it's time to install and load the plotly packages
```{r}
install.packages("plotly")                                        # Install plotly package
library("plotly")                                                 # Load plotly package

```
The important thing to remember here is that R is either not smart or it's very tired and needs very direct instructions. Just like me. 

We'll have to tell it which dataframe to look in and which axes will plot which column names.

Fig= plot_ly(x = data$column_name, y = data$other_column_name, z = data$column_to_be_heatmapped, type = "heatmap")

In this case, I want the Treatment column to be on the x axis and the Cytokine column to be on the Y. I want the D10 column which has my ratios to be my heatmap filling.
I'm going to call this plot output Fig.
```{r}
Fig= plot_ly(x = data$Treatment, y = data$Cytokine, z = data$D10, type = "heatmap")
  #set up your data
  
Fig
  #Print the heatmap to the viewer
```


